### 2025-04-23 Adipocyte_archetypes.Rmd
Added **Adipocyte_archetypes.Rmd** and updated **README**

---
title: "Adipocyte archetypes"
author: "Rasmus Rydbirk"
date: "'Sys.Date()'"
output: 
  html_document:
    toc: yes
    toc_float: yes
---

# Setup

```{r setup, message=FALSE}
library(archetypes)
library(magrittr)
library(dplyr)
library(qs)
library(scHelper) # github.com/rrydbirk/scHelper
```

# Identify archetypes

This is the code we used to identify archetypes. We set k = 3 based on cluster stability using *findCommunities* in **Conos** with parameter *test.stability = TRUE*. Please note, the results may not be exactly reproducible.

```{r}
con <- qread("con_adipocytes.qs", nthreads = 10)
emb <- con$embeddings$largeVis

res <- lapply(seq(10), \(x) {
  message(paste0("Iteration ",x))
  # Calculate archetypes
  tmp <- archetypes(data = emb, 3, verbose = F)

  # Right
  x <- 0.25
  a <- tmp$alphas %>% 
    as.data.frame() %>% 
    `rownames<-`(tmp$residuals %>% rownames()) %>% 
    filter(V1 < x, V2 > (1-x)) %>% 
    rownames()

  # Left
  x <- 0.30
  b <- tmp$alphas %>% 
    as.data.frame() %>% 
    `rownames<-`(tmp$residuals %>% rownames()) %>% 
    filter(V1 > (1-x), V2 < x) %>% 
    rownames()

  # Top
  x <- 0.25
  c <- tmp$alphas %>% 
    as.data.frame() %>% 
    `rownames<-`(tmp$residuals %>% rownames()) %>% 
    filter(V1 < x, V2 < x) %>% 
    rownames()

  out <- list(a,b,c)

  nn <- out %>% 
    sapply(\(x) unname(anno[x])) %>% 
    lapply(table) %>% 
    lapply(sort, T) %>% 
    lapply(names) %>% 
    sget(1)

  if (any(table(nn) > 1)) out <- NA else names(out) <- nn

  return(out)
}) %>% 
  .[!lapply(res, is.na) %>% sapply(any)]

anno.arc <- c("Adipocytes_RPSS23","Adipocytes_DGAT2","Adipocytes_CLSTN2") %>% 
  `names<-`(lapply(., \(x) sget(res, x) %>% 
                     Reduce(c, .) %>% 
                     table() %>% 
                     sort(T) %>% 
                     .[seq(2e3)] %>% 
                     names()), .) %>% 
  {lapply(seq(3), \(x) `names<-`(rep(names(.)[x], 2e3), .[[x]]))} %>% 
  Reduce(c, .) %>% 
  factor()

anno.arc %<>% factor(labels = c("CLSTN2_archetype", "DGAT2_archetype", "PRSS23_archetype"))

qsave(anno.arc, "anno_adipocytes_archetypes.qs")
```